### 2023-11-20 Manuscript/Draft2.Rmd
update manuscript

library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

knitr::opts_chunk$set(echo = FALSE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

@@

# Estimating Population Trends Under the Pressures of Climate Change

@@

```{r temp-scenarios, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="Mean trends of temperature data used in our MixFishSim population model."}
knitr::include_graphics(c("Images/TempScenarios.png"))
```

@@

Our analysis showed that both the design-based and model-based approaches are each capable of providing biomass estimates that track the true biomass trend. While VAST estimates tended to provide lower absolute error values compared to the stratified mean, especially when covariate information was included in the models, VAST Settings A and B also combined to produce over 30 biomass estimates with larger error values than the largest stratified mean error. These facts demonstrate the power of model-based estimates while also highlighting how sensitive they are to the choices made regarding settings and data used in analyses. Users often rely on traditional model-building processes and diagnostic techniques such as AIC to guide their choices, but successfully navigating the many decisions depends on the background of the given user and care should be taken in determining which diagnostic tools are used. For example, a recent simulation study showed that the existing diagnostic tools available in VAST sometimes guide users towards using settings that decrease the accuracy of the model **cite Chris C paper**. The value of model-based approaches such as VAST are clear- they allow users to overcome issues related to the input data to account for factors such as environmental change. However, the many consequential decisions that must be made in a given model and the lack of clear criteria for making these decisions makes it challenging to make these choices and even more difficult to determine whether the choices are correct.

@@

<div align="center">Table 1. Parameters used in all population models. SAW 1: [@nefsc1218], SAW 2: [@nefsc1311], SAW 3: not out yet??? </div>

@@

tabl <- "
|	Parameter	|	Description | Unit |Yellowtail | Cod | Haddock | Source
|:---	|:------- |:-- |:--- |:-- |:-- |:--
| $\\rho$ | Ford's growth coefficient | wk$^{-1}$  | 4.48 | 4.43 | 4.49 | [@thorson2020predicting] |
| $M$ | Natural Mortality| wk$^{-1}$  | 0.2064 | 0.2728 | 0.3340 | [@thorson2020predicting]|
| $F$ | Fishing Mortality| wk$^{-1}$  | 0.358 | 0.511 | 0.45 | SAW 1, 2, 3 |
| $W_R$ |Weight of fully recruited fish| kg | 0.39 | 2.95 | 1.12 | SAW 1, 2, 3 |
| $W_{R-1}$ | Weight of pre-recruit fish| kg | 0.13 | 0.39 | 0.19 | SAW 1, 2, 3 |
| $\\sigma^2$ | Variance in recruited fish | kg$^2$ | 0.55 | 0.55 | 0.55 |assumed|
| $\\lambda$ | Decay rate for movement | - | 0.7 | 0.7 | 0.7 |assumed|
| $Spwn_s$ | Spawning weeks for species $s$ | wk | 9-12 | 8-13 | 11-14 |SAW 1, 2, 3|
| $Rec_s$ | Recruitment weeks for species $s$ | wk | 9-12 | 8-13 | 11-14 |SAW 1, 2, 3|
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion

@@

# Create table to present the data]
Fig_5a <- kable(summary_tab_all[, -1],
      align = c("l", "l", "l", "l", "c", "c", "c"),
  caption = "Yellowtail Flounder absolute error results with all strata included in calculations. Row colors correspond to the same settings applied in different seasons.") %>%
  row_spec(0,bold = TRUE)%>% #bold column titles
  kable_styling(font_size = 10)  %>%
  pack_rows(tab_kable(), colnum = 1,
    index = table(fct_inorder(summary_tab_all$Scenario), useNA = "no"))%>%
  #color every 4th row a different color
  row_spec(seq(1,96,4), bold = T, color = "white", background = "black")%>%
  row_spec(seq(2,96,4), bold = T, color = "white", background = "#5C5E60")%>%
  row_spec(seq(3,96,4), bold = T, color = "white", background = "#A4A4A4")%>%
  row_spec(seq(4,96,4), bold = T, color = "white", background = "#D2D2D2")

@@

@@

>>>>

library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")
#library("embedr")
library("vembedr")
knitr::opts_chunk$set(echo = FALSE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

@@

# How might climate change affect indices of abundance? A simulation study starting point.

@@

@@

Our analysis showed that both the design-based and model-based approaches are each capable of providing biomass estimates that track the true biomass trend. While VAST estimates tended to provide lower absolute error values compared to the stratified mean, especially when covariate information was included in the models, VAST Settings A and B also combined to produce over 30 biomass estimates with larger error values than the largest stratified mean error. These facts demonstrate the power of model-based estimates while also highlighting how sensitive they are to the choices made regarding settings and data used in analyses. Users often rely on traditional model-building processes and diagnostic techniques such as AIC to guide their choices, but successfully navigating the many decisions depends on the background of the given user and care should be taken in determining which diagnostic tools are used. For example, a recent simulation study showed that the existing diagnostic tools available in VAST sometimes guide users towards using settings that decrease the accuracy of the model [@ChrisC]. The value of model-based approaches such as VAST are clear- they allow users to overcome issues related to the input data to account for factors such as environmental change. However, the many consequential decisions that must be made in a given model and the lack of clear criteria for making these decisions makes it challenging to make these choices and even more difficult to determine whether the choices are correct.

@@

<div align="center">Table 1. Parameters used in all population models. </div>

@@

tabl <- "
|	Parameter	|	Description | Unit |Yellowtail | Cod | Haddock | Source
|:---	|:------- |:-- |:--- |:-- |:-- |:--
| $\\rho$ | Ford's growth coefficient | wk$^{-1}$  | 4.48 | 4.43 | 4.49 | [@thorson2020predicting] |
| $M$ | Natural Mortality| wk$^{-1}$  | 0.2064 | 0.2728 | 0.3340 | [@thorson2020predicting]|
| $F$ | Fishing Mortality| wk$^{-1}$  | 0.358 | 0.511 | 0.45 | [@ManagTrack] |
| $W_R$ |Weight of fully recruited fish| kg | 0.39 | 2.95 | 1.12 | [@ManagTrack] |
| $W_{R-1}$ | Weight of pre-recruit fish| kg | 0.13 | 0.39 | 0.19 | [@ManagTrack] |
| $\\sigma^2$ | Variance in recruited fish | kg$^2$ | 0.55 | 0.55 | 0.55 |assumed|
| $\\lambda$ | Decay rate for movement | - | 0.7 | 0.7 | 0.7 |assumed|
| $Spwn_s$ | Spawning weeks for species $s$ | wk | 9-12 | 8-13 | 11-14 |[@ManagTrack]|
| $Rec_s$ | Recruitment weeks for species $s$ | wk | 9-12 | 8-13 | 11-14 |[@ManagTrack]|
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion

@@

# COLORs DEFINED BY ROW_SPEC ONLY WORK FOR PDF FORMAT
# Create table to present the data]
Fig_5a <- kable(summary_tab_all[, -1],
      align = c("l", "l", "l", "l", "c", "c", "c"),
  caption = "Yellowtail Flounder absolute error results with all strata included in calculations. Row colors correspond to the same settings applied in different seasons.") %>%
  row_spec(0,bold = TRUE)%>% #bold column titles
  kable_styling(font_size = 10)  %>%
  pack_rows(tab_kable(), colnum = 1,
    index = table(fct_inorder(summary_tab_all$Scenario), useNA = "no"))%>%
  #color every 4th row a different color
  row_spec(seq(1,96,4), bold = T, color = "white", background = "black")%>%
  row_spec(seq(2,96,4), bold = T, color = "white", background = "#5C5E60")%>%
  row_spec(seq(3,96,4), bold = T, color = "white", background = "#A4A4A4")%>%
  row_spec(seq(4,96,4), bold = T, color = "white", background = "#D2D2D2")

# COLORs DEFINED BY ROW_SPEC ONLY WORK FOR HTML FORMAT
# Fig_5a <- kable(summary_tab_all[, -1],
#       align = c("l", "l", "l", "l", "c", "c", "c"),
#   caption = "Yellowtail Flounder absolute error results with all strata included in calculations. Row colors correspond to the same settings applied in different seasons.") %>%
#   row_spec(0,bold = TRUE)%>% #bold column titles
#   kable_styling(font_size = 10)  %>%
#   pack_rows(tab_kable(), colnum = 1,
#     index = table(fct_inorder(summary_tab_all$Scenario), useNA = "no"))

@@

#COLOR INDIVIDUAL ROWS BY CATEGORY
#define colors
# 
#  COLORS DEFINED BY ROW_SPEC ONLY WORK FOR PDF FORMAT

@@

#COLORS BELOW DEFINED BY ROW_SPEC ARE FOR HTML OUTPUT

# Fig_5b <- kable(summary_tab_exclude[, -1],
#       align = c("l", "l", "l", "l", "c", "c", "c"),
#   caption = "Yellowtail Flounder absoloute error results with certain strata excluded from calculations. Row colors correspond to the same settings applied in different seasons.") %>%
#   row_spec(0,bold = TRUE)%>% #bold column titles
#   kable_styling(font_size = 10)  %>%
#   pack_rows(tab_kable(), colnum = 1,
#     index = table(fct_inorder(summary_tab_exclude$Scenario), useNA = "no"))

### 2023-11-20 Manuscript/Draft2.tex
update manuscript

\hypertarget{estimating-population-trends-under-the-pressures-of-climate-change}{%
\section{Estimating Population Trends Under the Pressures of Climate Change}\label{estimating-population-trends-under-the-pressures-of-climate-change}}

@@

{\centering \includegraphics[width=0.95\linewidth]{Images/TempScenarios}

\caption{Mean trends of temperature data used in our MixFishSim population model.}\label{fig:temp-scenarios}
\end{figure}

@@

Our analysis showed that both the design-based and model-based approaches are each capable of providing biomass estimates that track the true biomass trend. While VAST estimates tended to provide lower absolute error values compared to the stratified mean, especially when covariate information was included in the models, VAST Settings A and B also combined to produce over 30 biomass estimates with larger error values than the largest stratified mean error. These facts demonstrate the power of model-based estimates while also highlighting how sensitive they are to the choices made regarding settings and data used in analyses. Users often rely on traditional model-building processes and diagnostic techniques such as AIC to guide their choices, but successfully navigating the many decisions depends on the background of the given user and care should be taken in determining which diagnostic tools are used. For example, a recent simulation study showed that the existing diagnostic tools available in VAST sometimes guide users towards using settings that decrease the accuracy of the model \textbf{cite Chris C paper}. The value of model-based approaches such as VAST are clear- they allow users to overcome issues related to the input data to account for factors such as environmental change. However, the many consequential decisions that must be made in a given model and the lack of clear criteria for making these decisions makes it challenging to make these choices and even more difficult to determine whether the choices are correct.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-nefsc1218}{}}%
NEFSC. 2012. {54th Northeast Regional Stock Assessment Workshop (54th SAW) Assessment Report}. {Northeast Fish. Sci. Cent. Ref. Doc. 12-18; 604 p. Available from: National Marine Fisheries Service, 166 Water Street, Woods Hole, MA 02543-1026, or online at http://www.nefsc.noaa.gov/nefsc/publications/}.

\leavevmode\vadjust pre{\hypertarget{ref-nefsc1311}{}}%
NEFSC. 2013. {55th Northeast Regional Stock Assessment Workshop (55th SAW) Assessment Report}. {Northeast Fish. Sci. Cent. Ref. Doc. 13-11; 849 p. Available from: National Marine Fisheries Service, 166 Water Street, Woods Hole, MA 02543-1026, or online at http://www.nefsc.noaa.gov/nefsc/publications/}.

@@

Table 1. Parameters used in all population models. SAW 1: (NEFSC 2012), SAW 2: (NEFSC 2013), SAW 3: not out yet???

\begin{longtable}[]{@{}lllllll@{}}
\toprule
Parameter & Description & Unit & Yellowtail & Cod & Haddock & Source \\
\midrule
\endhead
\(\rho\) & Ford's growth coefficient & wk\(^{-1}\) & 4.48 & 4.43 & 4.49 & (Thorson 2020) \\
\(M\) & Natural Mortality & wk\(^{-1}\) & 0.2064 & 0.2728 & 0.3340 & (Thorson 2020) \\
\(F\) & Fishing Mortality & wk\(^{-1}\) & 0.358 & 0.511 & 0.45 & SAW 1, 2, 3 \\
\(W_R\) & Weight of fully recruited fish & kg & 0.39 & 2.95 & 1.12 & SAW 1, 2, 3 \\
\(W_{R-1}\) & Weight of pre-recruit fish & kg & 0.13 & 0.39 & 0.19 & SAW 1, 2, 3 \\
\(\sigma^2\) & Variance in recruited fish & kg\(^2\) & 0.55 & 0.55 & 0.55 & assumed \\
\(\lambda\) & Decay rate for movement & - & 0.7 & 0.7 & 0.7 & assumed \\
\(Spwn_s\) & Spawning weeks for species \(s\) & wk & 9-12 & 8-13 & 11-14 & SAW 1, 2, 3 \\
\(Rec_s\) & Recruitment weeks for species \(s\) & wk & 9-12 & 8-13 & 11-14 & SAW 1, 2, 3 \\
\bottomrule
\end{longtable}

>>>>

\hypertarget{how-might-climate-change-affect-indices-of-abundance-a-simulation-study-starting-point.}{%
\section{How might climate change affect indices of abundance? A simulation study starting point.}\label{how-might-climate-change-affect-indices-of-abundance-a-simulation-study-starting-point.}}

@@

@@

Our analysis showed that both the design-based and model-based approaches are each capable of providing biomass estimates that track the true biomass trend. While VAST estimates tended to provide lower absolute error values compared to the stratified mean, especially when covariate information was included in the models, VAST Settings A and B also combined to produce over 30 biomass estimates with larger error values than the largest stratified mean error. These facts demonstrate the power of model-based estimates while also highlighting how sensitive they are to the choices made regarding settings and data used in analyses. Users often rely on traditional model-building processes and diagnostic techniques such as AIC to guide their choices, but successfully navigating the many decisions depends on the background of the given user and care should be taken in determining which diagnostic tools are used. For example, a recent simulation study showed that the existing diagnostic tools available in VAST sometimes guide users towards using settings that decrease the accuracy of the model (Christopher Cacciapaglia 2023). The value of model-based approaches such as VAST are clear- they allow users to overcome issues related to the input data to account for factors such as environmental change. However, the many consequential decisions that must be made in a given model and the lack of clear criteria for making these decisions makes it challenging to make these choices and even more difficult to determine whether the choices are correct.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ChrisC}{}}%
Christopher Cacciapaglia, C.F.A., Elizabeth N. Brooks. 2023. Developing workflow and diagnostics for model selection of a vector autoregressive spatiotemporal (VAST) model. Fisheries Research \textbf{In Press}. Elsevier.

@@

@@

Table 1. Parameters used in all population models.

\begin{longtable}[]{@{}lllllll@{}}
\toprule
Parameter & Description & Unit & Yellowtail & Cod & Haddock & Source \\
\midrule
\endhead
\(\rho\) & Ford's growth coefficient & wk\(^{-1}\) & 4.48 & 4.43 & 4.49 & (Thorson 2020) \\
\(M\) & Natural Mortality & wk\(^{-1}\) & 0.2064 & 0.2728 & 0.3340 & (Thorson 2020) \\
\(F\) & Fishing Mortality & wk\(^{-1}\) & 0.358 & 0.511 & 0.45 & ((NEFSC) 2022) \\
\(W_R\) & Weight of fully recruited fish & kg & 0.39 & 2.95 & 1.12 & ((NEFSC) 2022) \\
\(W_{R-1}\) & Weight of pre-recruit fish & kg & 0.13 & 0.39 & 0.19 & ((NEFSC) 2022) \\
\(\sigma^2\) & Variance in recruited fish & kg\(^2\) & 0.55 & 0.55 & 0.55 & assumed \\
\(\lambda\) & Decay rate for movement & - & 0.7 & 0.7 & 0.7 & assumed \\
\(Spwn_s\) & Spawning weeks for species \(s\) & wk & 9-12 & 8-13 & 11-14 & ((NEFSC) 2022) \\
\(Rec_s\) & Recruitment weeks for species \(s\) & wk & 9-12 & 8-13 & 11-14 & ((NEFSC) 2022) \\
\bottomrule
\end{longtable}
